Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-20 doc/Paper/main.tex
my changes

\author[a,1]{Andrina Nicola,}
\author[b]{David Alonso,}
\author[c]{An\v{z}e Slosar,}
\author[d]{Humna Awan,}
\author[d]{Adam Broussard,}
\author[b]{Zahra Gomes,}
\author[d]{Eric Gawiser,}
\author[x]{Rachel Mandelbaum,}
\author[x]{Hironao Miyatake,}
\author[e]{Jeffrey A. Newman,}
\author[f,g]{Javier S\'anchez,}
\author[x]{Ignacio Sevilla,}
\author[c,h]{Sarah Skinner,}
\author[x]{Erica Wagoner}

\affiliation[a]{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton NJ 08544-0010, USA}
\affiliation[b]{Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom}
\affiliation[c]{Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
\affiliation[d]{Rutgers University, Physics \& Astronomy Department, Piscataway, NJ 08854, USA}
\affiliation[e]{Department of Physics and Astronomy and PITT PACC, University of Pittsburgh, Pittsburgh, PA, 15260, USA}
\affiliation[f]{Fermi National Accelerator Laboratory, Batavia, IL, 60510, USA}
\affiliation[g]{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\affiliation[h]{Department of Physics, Missouri University of Science and Technology, Rolla, MO 65401, USA}

@@

The past two decades have seen a revolution in our understanding of the universe and its constituents. While observations of the cosmic microwave background play a pivotal role in anchoring the standard  cosmological model \cite{Planck:2018}, they are not a direct probe of the physics of the low redshift universe. In order to directly characterize properties of dark energy and modified gravity, we need to measure the expansion history and growth  \cite{1903.12016}. At the moment, this is only possible with the optical large scale structure surveys.

Over a decade ago, the Dark Energy Task Force has divided the evolution of optical experiments into approximate stages \cite{0609591}. The current Stage III experiments, including spectroscopic surveys such as eBOSS \cite{1707.09322,Zhu:2018,1712.08064,1801.02689,1801.02891} and VIPERS \cite{1611.07048,1612.05645,1708.00026} and photometric surveys such as DES \cite{1708.01531,1708.01530,1708.01536}, the Hyper Suprime-Cam \cite{1704.05858,2019PASJ...71...43H,1906.06041} and the Kilo-Degree Survey / VIKING-450 \cite{1902.11265,1812.06077,Joudaki:2019} are approaching completion and the field is preparing for the Stage IV surveys, such as DESI \cite{1611.00036,1907.10688} and LSST \cite{0912.0201,1809.01669} as well as major satellite missions like Euclid \cite{1606.00180} and WFIRST \cite{1904.01174}.

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer source redshift distributions is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as scattered light from bright objects, errors in star-galaxy separation, reddening by milky way dust and others will imprint spurious fluctuations on the observed galaxy density field. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density fields. Therefore, Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

The community is well aware of these challenges lying ahead. There are two main approaches to preparing for the arrival of data. On one hand, we are building sophisticated, realistic mock data sets, such as LSST DESC Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are (re-)analyzing existing precursor data to test-drive our methods and codes in a realistic environment. In this paper, we focus on the latter approach and perform an analysis of photometric galaxy clustering using the first data release (DR1) of the Hyper Suprime-Cam Subaru Strategic Program (HSC-SSP) \cite{2018PASJ...70S...8A}. This dataset bears significant similarities to the data expected from LSST, both in terms of survey depth and photometric bands as well as primary data reduction and catalog generation. Specifically, HSC covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes. In this work, we focus on a magnitude-limited galaxy sample with limiting magnitude for 5-$\sigma$ source detection of $i<24.5$, which is similar to the expected magnitude cut to be used for LSST of $i<25.1$ after 1 year \cite{1809.01669}. Finally, the primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is therefore a perfect proxy for future LSST data.

In this work, we present an analysis of photometric galaxy clustering using HSC DR1 data. We split the data into four tomographic redshift bins and measure both auto- and cross power spectra, correcting for observational systematics using mode deprojection. We then use the halo model coupled with a halo occupation distribution to fit the data and derive constraints on astrophysical parameters, marginalizing over photometric redshift uncertainties. To ensure the stability of our results, we perform a suite of complementary analyses and find our constraints to be robust to photometric redshift uncertainties. We additionally perform extended analyses in which we compute constraints on magnification in our sample and allow for variations in the $\Lambda$CDM cosmological parameters, respectively. Finally, we use our data to derive a simple fitting function for the linear galaxy bias as a function of redshift and limiting magnitude of the sample.

In our analysis, we make several simplifying approximations; in particular we employ a simple semi-analytical halo model to model the data, which will most certainly not be sophisticated enough for the full sky LSST analysis. LSST data will likely require  a combination of  bias expansion approaches \cite{0902.0991,1402.5916,1611.09787,1910.07097} and power spectrum emulators \cite{1804.05865,1705.03388}. However, given the level of statistical precision of the data at hand, we find that a simple halo model suffices, and thus use we use it as a ``good enough'' substitute for the final modeling.

@@

  \begin{table}
  \centering
  \begin{tabular}{|l|r|c|l|}
   \hline
   {\bf Field name} & $N_{\rm gal}$ & {\bf Area} (deg$^2$) & $f_{\rm sky}$ \\
   \hline
   GAMA09H  & 1,697,713 & 14.5 & $3.5\times10^{-4}$ \\
   GAMA15H  & 1,695,364 & 15.1 & $3.7\times10^{-4}$ \\
   HECTOMAP &   639,970 &  5.1 & $1.2\times10^{-4}$ \\
   VVDS     & 2,340,965 & 20.6 & $5.0\times10^{-4}$ \\
   WIDE12H  & 1,220,816 & 11.6 & $2.8\times10^{-4}$ \\
   XMMLSS   & 2,139,629 & 20.8 & $5.0\times10^{-4}$ \\
   \hline
   Total    & 9734457 & 87.7 & $2.1\times10^{-3}$ \\
   \hline
  \end{tabular}
  \caption{Summary of the 6 different fields used in our analysis. The second column lists the number of galaxies in the DR1 catalog passing the cuts in Table \ref{tab:cuts_summary}. The third and fourth columns show the area and corresponding sky fraction covered by each field. The area was calculated as the sum of pixel areas allowed by the sky mask described in Section \ref{ssec:methods.mask}. We note that HECTOMAP field is not included in our analysis due to its smaller area.} \label{tab:field_summary}
  \end{table}

@@

In this work, we use data from the first data release of the Hyper Suprime-Cam Subaru Strategic Program (HSC DR1 hereon)\footnote{\texttt{https://hsc-release.mtk.nao.ac.jp}}. The release is extensively documented in \cite{2018PASJ...70S...8A}, and we only provide the details of the galaxy sample and associated data used for our clustering analysis.

  HSC-SSP is a photometric galaxy survey that has been awarded 300 nights on the Subaru Telescope starting in 2014. DR1 includes data from 61.5 nights observed to three different depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degrees to $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 7 fields  of areas varying between 5 and 20 square degrees. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than the median seeing that will be achieved by LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and are available to the community through a publicly available database.

@@

      \item {\bf Observing conditions}: the HSC DR1 provides metadata for each filter exposure, containing information about a number of observing conditions. We make coadded maps of these following a procedure similar to that described in \cite{2016ApJS..226...24L}. For each pixel in our map we gather all exposures that fully or partially overlap with it. For each quantity $Q$ in filter $f$, this allows us to build a list of values of $Q$ for each exposure\footnote{It is worth noting that we omit any exposure from CCD 9, since it was found to yield unreliable measurements and was never used in the HSC coadd images.}. We then compute the weighted mean of these values and assign the result to the pixel. The weights used to coadd different exposures consist of the product of the area overlap between pixel and exposure and an approximation of the weights used by the HSC pipeline to produce coadded images \cite{2018PASJ...70S...5B}. Since we do not have direct access to the latter, we used, as coadd weights, the inverse of the sky level of each exposure, which should be close to inverse-variance weighting. Following this procedure we produce maps of the following quantities: airmass, CCD temperature, seeing, point-spread function (PSF) ellipticity, exposure time, sky count level, root-mean square deviation of the sky count and number of visits.

@@

We have also investigated the impact of a secondary mode in the redshift distribution for the first tomographic bin. This secondary peak contains $\sim$5\% of galaxies and peaks around $z\sim 3$ (hence cannot be seen in Figure \ref{fig:nzs}). To investigate the impact of this mode, we manually excise it from redshift distribution and calculate the resulting change in $\Delta \chi^2\sim 0.05$. This indicates that the effect of this bump is to small to be statistically significant. We therefore do not need to worry whether it is real or an artifact of photometric redshift inference.

>>>>

\author[a,1]{Andrina Nicola,}
\author[b]{David Alonso,}
\author[c]{An\v{z}e Slosar,}
\author[d,e]{Javier S\'anchez,}
\author[f]{Humna Awan,}
\author[f]{Adam Broussard,}
\author[b]{Zahra Gomes,}
\author[f]{Eric Gawiser,}
\author[x]{Rachel Mandelbaum,}
\author[x]{Hironao Miyatake,}
\author[g]{Jeffrey A. Newman,}
\author[x]{Ignacio Sevilla,}
\author[c,h]{Sarah Skinner,}
\author[x]{Erica Wagoner}

\affiliation[a]{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton NJ 08544-0010, USA}
\affiliation[b]{Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom}
\affiliation[c]{Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA}
\affiliation[d]{Fermi National Accelerator Laboratory, Batavia, IL, 60510, USA}
\affiliation[e]{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\affiliation[f]{Rutgers University, Physics \& Astronomy Department, Piscataway, NJ 08854, USA}
\affiliation[g]{Department of Physics and Astronomy and PITT PACC, University of Pittsburgh, Pittsburgh, PA, 15260, USA}
\affiliation[h]{Department of Physics, Missouri University of Science and Technology, Rolla, MO 65401, USA}

@@

The past two decades have seen a revolution in our understanding of the Universe and its constituents. While observations of the cosmic microwave background (CMB) play a pivotal role \cite{Planck:2018}, they only indirectly probe the physics of the low-redshift Universe, in particular the expansion history and growth, which are uniquely sensitive to properties of dark energy and modified gravity \cite{1903.12016}.

In studying the late-time Universe, surveys of the large scale structure play an increasingly important role. Over a decade ago, the Dark Energy Task Force divided the evolution of optical experiments into approximate stages \cite{0609591}. The current Stage III experiments, including spectroscopic surveys such as eBOSS \cite{1707.09322,Zhu:2018,1712.08064,1801.02689,1801.02891} and VIPERS \cite{1611.07048,1612.05645,1708.00026} and photometric surveys such as DES \cite{1708.01531,1708.01530,1708.01536}, the Hyper Suprime-Cam \cite{1704.05858,2019PASJ...71...43H,1906.06041} and the Kilo-Degree Survey / VIKING-450 \cite{1902.11265,1812.06077,Joudaki:2019} are approaching completion, and the field is preparing for the Stage IV surveys, such as the ground-based DESI \cite{1611.00036,1907.10688} and LSST \cite{0912.0201,1809.01669}, as well as major satellite missions like Euclid \cite{1606.00180} and WFIRST \cite{1904.01174}.

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function (PSF) with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer source redshift distributions is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as scattered light from bright objects, errors in star-galaxy separation, reddening by Milky Way dust, and varying observing conditions, will imprint spurious fluctuations on the observed galaxy density field. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density field. Therefore, Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

The community is well aware of these challenges lying ahead. There are two main approaches to preparing for the arrival of data. On one hand, we are building sophisticated, realistic mock data sets, such as LSST DESC Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are (re-)analyzing existing precursor data to test-drive our methods and codes in a realistic environment. In this paper, we focus on the latter approach and perform an analysis of photometric galaxy clustering using the first data release (DR1) of the Hyper Suprime-Cam Subaru Strategic Program (HSC-SSP) \cite{2018PASJ...70S...8A}. This dataset bears significant similarities to the data expected from LSST, both in terms of survey depth and photometric bands as well as primary data reduction and catalog generation. Specifically, HSC covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes, which are unlikely to play a major role in cosmological data analysis. In this work, we focus on a magnitude-limited galaxy sample with limiting magnitude of $i<24.5$, which is similar to the expected 5-$\sigma$ detection limit for LSST after 1 year \cite{1809.01669} ($i<25.1$). Finally, the primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage, employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is therefore a perfect proxy for future LSST data. This analysis has the additional attraction that no other photometric galaxy clustering study has been carried out with these data.

In this work, we present an analysis of photometric galaxy clustering using HSC DR1 data. We split the data into four tomographic redshift bins and measure both auto- and cross power spectra between all bins, correcting for observational systematics using mode deprojection. We then use the halo model coupled with a halo occupation distribution to fit the data and derive constraints on astrophysical parameters, marginalizing over photometric redshift uncertainties. To ensure the stability of our results, we perform a suite of complementary analyses and find our constraints to be robust to photometric redshift uncertainties. We additionally perform extended analyses in which we compute constraints on magnification in our sample and allow for variations in the $\Lambda$CDM cosmological parameters, respectively. Finally, we use our data to derive a simple fitting function for the linear galaxy bias as a function of redshift and limiting magnitude of the sample. In our analysis, we make several simplifying approximations; in particular we employ a simple semi-analytical halo model to model the data, which will most certainly not be sophisticated enough for the full sky LSST analysis. However, we find that it suffices for the statistical precision of the data at hand, and we view the use of the halo model as a ``good enough'' substitute for the final modeling, which will likely include bias expansion approaches \cite{0902.0991,1402.5916,1611.09787} and power spectrum emulators \cite{1804.05865,1705.03388}.

@@

  \begin{table}
  \centering
  \begin{tabular}{|l|r|c|l|}
   \hline
   {\bf Field name} & $N_{\rm gal}$ & {\bf Area} (deg$^2$) & $f_{\rm sky}$ \\
   \hline
   GAMA09H  & 1,697,713 & 14.5 & $3.5\times10^{-4}$ \\
   GAMA15H  & 1,695,364 & 15.1 & $3.7\times10^{-4}$ \\
   HECTOMAP &   639,970 &  5.1 & $1.2\times10^{-4}$ \\
   VVDS     & 2,340,965 & 20.6 & $5.0\times10^{-4}$ \\
   WIDE12H  & 1,220,816 & 11.6 & $2.8\times10^{-4}$ \\
   XMMLSS   & 2,139,629 & 20.8 & $5.0\times10^{-4}$ \\
   \hline
   Total    & 9734457 & 87.7 & $2.1\times10^{-3}$ \\
   \hline
  \end{tabular}
  \caption{Summary of the 6 different fields used in our analysis. The second column lists the number of galaxies in the DR1 catalog passing the cuts in Table \ref{tab:cuts_summary}. The third and fourth columns show the area and corresponding sky fraction covered by each field. The area was calculated as the sum of pixel areas allowed by the sky mask described in Section \ref{ssec:methods.mask}. \an{We don't use HECTOMAP in the fitting. Shall we even show it here?}} \label{tab:field_summary}
  \end{table}

@@

  In this work, we use data from the first data release of the Hyper Suprime-Cam Subaru Strategic Program (HSC DR1 hereon)\footnote{\url{https://hsc-release.mtk.nao.ac.jp}}. The release is extensively documented in \cite{2018PASJ...70S...8A}, and we only provide the details of the galaxy sample and associated data used for our clustering analysis.

  HSC-SSP is a photometric galaxy survey that has been awarded 300 nights on the Subaru Telescope starting in 2014. DR1 includes data from 61.5 nights observed to three different depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degrees to $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 6 fields  of areas varying between 5 and 20 square degrees. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than the median seeing that will be achieved by LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and are available to the community through a public database.

@@

      \item {\bf Observing conditions}: the HSC DR1 provides metadata for each filter exposure, containing information about a number of observing conditions. We make coadded maps of these following a procedure similar to that described in \cite{2016ApJS..226...24L}. For each pixel in our map we gather all exposures that fully or partially overlap with it. For each quantity $Q$ in filter $f$, this allows us to build a list of values of $Q$ for each exposure\footnote{It is worth noting that we omit any exposure from CCD 9, since it was found to yield unreliable measurements and was never used in the HSC coadd images.}. We then compute the weighted mean of these values and assign the result to the pixel. The weights used to coadd different exposures consist of the product of the area overlap between pixel and exposure and an approximation of the weights used by the HSC pipeline to produce coadded images \cite{2018PASJ...70S...5B}. Since we do not have direct access to the latter, we used, as coadd weights, the inverse of the sky level of each exposure, which should be close to inverse-variance weighting. Following this procedure we produce maps of the following quantities: airmass, CCD temperature, seeing, PSF ellipticity, exposure time, sky count level, root-mean square deviation of the sky count and number of visits.

@@

We have also investigated the impact of a secondary mode in the redshift distribution for the first tomographic bin. This secondary peak contains 5\% of galaxies and peaks around $z\sim 2.5$ (hence cannot be seen in Figure \ref{fig:nzs}). To investigate the impact of this mode, we manually excise it from redshift distribution and calculate the resulting change in $\Delta \chi^2\sim 0.1$. This indicates that the effect of this bump is to small to be statistically significant. We therefore do not need to worry whether it is real or an artifact of photometric redshift inference.  \as{Fix numbers once Sarah gets back}
